HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-04-27 docs/HLL_exp_bad/main.tex
rm 'he' for attacker

\begin{itemize}
\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows him to submit an item and learn the hypothetical effect of inserting it into the sketch. He is then able to determine if inserting this item \textit{would} increase the cardinality estimation.

%The adversary is able to test the hypothetical effect of the insertion of an item on the cardinality estimation of the sketch. This can achieved via a \textit{test oracle} provided by the sketch owner, which allows an user to test the effect of an insertion on the cardinality estimate but without actually inserting the item, or if the attacker would gain access to a black-box replica of the sketch in use, i.e., buying the same switch and testing in the lab. %black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adverary can test the effect of an insertion without making the insertion!s}
\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. \mathilde{Would it make sense to add a sentence "This scenario would apply to~\cite{portscanhll}."?}
\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use. Additionally, he can access and interact with the sketch via an API provided by the sketch owner, allowing him to insert an item or ask for the HLL's cardinality estimate at any point in time.
\kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch. Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate. Again, an S4 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack, but it can compute these values for itself (assuming its attack proceeds without input from other users being added to the HLL sketch in parallel with the attack). \mathilde{I am not sure about the last sentence. In the \textit{insider} model of \cite{cardestprivacy} the attacker owns the sketch so has access to both statistics (cardinality estimation) and counters in real time. So if he receives data coming from another user / notices an update in the counters, he can check reiterate over the discarded items and check whether he can retain some more? I know the goal of \cite{cardestprivacy} is different and there it makes sense to consider the scenario where the owner is malicious but not necessarily for us.}
\end{itemize}

@@

Furthermore, we show that although \cite{cardestprivacy} presents S3 and S4 as detached scenarios, the attacker with the capabilities of S3 can obtain the knowledge of S4. By analyzing which inputs increase the cardinality estimation, an attacker can easily recover the contents of each bucket in the sketch. To guess the estimator of a targeted bucket, the attacker gradually inserts items with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$, until the cardinality estimation of the sketch increases. Once a change occurs, the attacker knows that the maximum number of leading zeros observed in that bucket just got updated to a value that can be recovered by looking at the last inserted item. The attacker can determine all estimators with $O(2^{n + \ell})$ pre-computations and $O(\ell \cdot 2^{n})$ insert and cardinality estimation queries. The fuller the sketch the most work the attacker has to provide.

>>>>

\begin{itemize}
\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows him to submit an item and learn the hypothetical effect of inserting it into the sketch. It is then able to determine if inserting this item \textit{would} increase the cardinality estimation.

%The adversary is able to test the hypothetical effect of the insertion of an item on the cardinality estimation of the sketch. This can achieved via a \textit{test oracle} provided by the sketch owner, which allows an user to test the effect of an insertion on the cardinality estimate but without actually inserting the item, or if the attacker would gain access to a black-box replica of the sketch in use, i.e., buying the same switch and testing in the lab. %black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adverary can test the effect of an insertion without making the insertion!s}
\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. \mathilde{Would it make sense to add a sentence "This scenario would apply to~\cite{portscanhll}."?}
\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing him to insert an item or ask for the HLL's cardinality estimate at any point in time.
\kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch. Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate. Again, an S4 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack, but it can compute these values for itself (assuming its attack proceeds without input from other users being added to the HLL sketch in parallel with the attack). \mathilde{I am not sure about the last sentence. In the \textit{insider} model of \cite{cardestprivacy} the attacker owns the sketch so has access to both statistics (cardinality estimation) and counters in real time. So if it receives data coming from another user / notices an update in the counters, it can check reiterate over the discarded items and check whether it can retain some more? I know the goal of \cite{cardestprivacy} is different and there it makes sense to consider the scenario where the owner is malicious but not necessarily for us.}
\end{itemize}

@@

Furthermore, we show that although \cite{cardestprivacy} presents S3 and S4 as detached scenarios, the attacker with the capabilities of S3 can obtain the knowledge of S4. By analyzing which inputs increase the cardinality estimation, an attacker can easily recover the contents of each bucket in the sketch. To guess the estimator of a targeted bucket, the attacker gradually inserts items with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$, until the cardinality estimation of the sketch increases. Once a change occurs, the attacker knows that the maximum number of leading zeros observed in that bucket just got updated to a value that can be recovered by looking at the last inserted item. The attacker can determine all estimators with $O(2^{n + \ell})$ pre-computations and $O(\ell \cdot 2^{n})$ insert and cardinality estimation queries. \mathilde{todo: add sentence for The fuller the sketch the most work the attacker has to provide.}
